Using Planning to Generate Natural-Language Sentences:\\ Experiences and Challenges

### 2009-12-11 trunk/Papers/crisp-ff-10/crisp-ff.tex
title, structure, ff sec outline modifications

\section{Experiences with FF}
\label{sec:crisp-ff}

@@

\begin{enumerate}
\item Enforced hillclimbing: Initial state of example problem has a
  relaxed graph of length one because ``expressed'' and ``subst''
  goals are achieved, and newly introduced violations of ``subst'' and
  ``distractor'' are not noticed. As a consequence, EHC on the second
  state spends all of its time in breadth-first search.
\item Goal agenda: For some reason that I don't understand in detail,
  the goal agenda heuristic can choose that the ``subst'' goal must be
  reached before even the ``expressed'' goal. This leads the planner
  to generating some arbitrary sentence. Because the example grammar
  is so small, this leads to a state from which the ``expressed'' goal
  can no longer be achieved. (A realistic big grammar should be able
  to recover from this by introducing a conjunction or relative clause
  or some such -- but of course, this would still lead to a stupid
  sentence.)
\item No helpful actions in best-first search: After the initial
  failure of the goal agenda, FF fell back to best-first search, but
  didn't use the helpful actions heuristic in this search.
\item Tautology detection in reachability analysis: The old FF
  performed a pairwise comparison of the ground literals in the
  instantiated goal to remove tautologies and obvious contradictions.
  In the sentence generation problems, the instantiated goal can
  contain lots of literals, so this takes a long time.  Removing the
  check helped a lot.  Another way to fix this problem would have been
  to sort the list of ground literals appropriately and then only
  compare neighboring literals.
\item Unnecessary computation of hard action template data structure:
  FF computed a big data structure that was only useful for hard
  action templates, although the sentence generation planning problems
  only contain easy action templates.  Because the universe can be
  relatively big, this again wasted a lot of time.
\end{enumerate}

>>>>

\section{Why FF off-the-shelf doesn't work, and how to fix it (partially)}
\label{sec:crisp-ff}

\subsection{Pre-Processing}
\label{sec:crisp-ff:preprocess}

@@

\begin{enumerate}
\item Tautology detection in reachability analysis: The old FF
  performed a pairwise comparison of the ground literals in the
  instantiated goal to remove tautologies and obvious contradictions.
  In the sentence generation problems, the instantiated goal can
  contain lots of literals, so this takes a long time.  FIX: Removing
  the check helped a lot.  Another way to fix this problem would have
  been to sort the list of ground literals appropriately and then only
  compare neighboring literals.
\item Unnecessary computation of hard action template data structure:
  FF computed a big data structure that was only useful for hard
  action templates, although the sentence generation planning problems
  only contain easy action templates.  Because the universe can be
  relatively big, this again wasted a lot of time. FIX: just add this
  tuiny bit of intelligence to FF to not do the data structure if it
  is not needed. .. interesting that no-one had this problem before.
\end{enumerate}

\subsection{Heuristics and Search}
\label{sec:crisp-ff:preprocess:heuristics}

\begin{enumerate}
\item Goal agenda: For some reason that I don't understand in detail,
  the goal agenda heuristic chooses that the ``subst'' goal must be
  reached before even the ``expressed'' goal. This leads the planner
  to generating some arbitrary sentence. Because the example grammar
  is so small, this leads to a state from which the ``expressed'' goal
  can no longer be achieved. (A realistic big grammar should be able
  to recover from this by introducing a conjunction or relative clause
  or some such -- but of course, this would still lead to a stupid
  sentence.) FIX: turn goal agenda off. alternatively, develop
  appropriate goal ordering techniques.
\item Subst and distractor atoms have a weird lifecycle: They start as
  false, then get made true at some point, and then false again.
  After they have been made false, they can never become true
  again. FIX: no fix just yet. highly non-trivial challenge to
  generation of h fns. {\bf (QUESTION: does lama do better? do any of
    the known admissible heuristics do better?)}
\item As a consequence, Enforced hillclimbing: Initial state of
  example problem has a relaxed graph of length one because
  ``expressed'' and ``subst'' goals are achieved, and newly introduced
  violations of ``subst'' and ``distractor'' are not noticed. As a
  consequence, EHC on the second state spends all of its time in
  breadth-first search. FIX: don't use EHC.
\item The only thing that works well is helpful actions. However,
  these didn't get around to be used due to no helpful actions in
  best-first search: After failure of EHC, FF falls back to best-first
  search, but doesn't use the helpful actions heuristic in this
  search. FIX: simply go straight to BFS with helpful actions.
\end{enumerate}

### 2009-12-11 trunk/Papers/crisp-ff-10/main.tex
some initial stuff
title, structure, ff sec outline modifications

@@

\begin{document}
\title{Challenge: Using Planning to Generate Natural-Language
  Sentences}
\author{J\"org Hoffmann \\ LORIA \\ joerg.hoffmann@loria.fr
\And
Alexander Koller \\ Saarland University \\ koller@mmci.uni-saarland.de
\And
Ronald Petrick \\ University of Edinburgh \\ rpetrick@inf.ed.ac.uk}
\maketitle

\input abstract
\input introduction
\input crisp
\input crisp-ff
\input conclusion

>>>>

\usepackage{graphicx}

@@

\begin{document}
\title{Using Planning to Generate Natural-Language Sentences:\\
  Experiences and Challenges}

\author{
Alexander Koller \\ Saarland University, Germany \\ koller@mmci.uni-saarland.de
\And
J\"org Hoffmann \\ INRIA, France \\ joerg.hoffmann@inria.fr
\And
Ronald Petrick \\ University of Edinburgh, UK \\ rpetrick@inf.ed.ac.uk
}

\maketitle

\input abstract
\input introduction
\input crisp
\input crisp-ff
\input experiments
\input conclusion
